Toy Transit Surveys

### 2017-08-30 memos/workdir/toy_analytic_surveys.tex
update memo with clearer correction factor.
also remove obnoxious TeX extras from repo

Finally, binarity may affect transit surveys through actual astrophysics.
For the relatively rare case [Duquennoy \& Mayor 1991] of close 
binaries, dynamical stability imposes a limited range of periods in which 
planets may orbit [e.g., Holman \& Weigert 
1999, Li et al 2017], the effects of which have been observed [Kraus et al 
2016].
Also, the formation rates of planets, and thus their intrinsic occurrence 
rates, may be influenced by properties of circumprimary and/or circumbinary 
disks.
Although the observations are still ongoing, early results seem to indicate 
that once binary companions are $\gtrsim 100\,{\rm AU}$ apart, binarity does 
not decrease planet occurrence rates [CITE], and may even increase them for 
giant planets [Ngo et al 2017].

@@

This is almost simply asking ``what is the relative difference between the 
occurrence rates derived by Astronomers D and A for planets of radius $R_p$?''
However, in the more realistic case, Astronomer A also has derived a 
completeness, which we assume is the same as for Astronomer D in the single 
star case.
So Astronomer A now misclassifies planetary radii, and miscounts the total 
number of stars, but knows his completeness for single stars.
Astronomer D corrects all these errors.
Then
\begin{equation}
\Gamma_{{\rm A}, R_p} = \frac{N_{\rm det,s}/(f_{s,g} f_{s,c})}{N_s + N_d},
\end{equation}
and the relative difference between the two occurrence rates is

\begin{align}
\left|\frac{\Gamma_{{\rm A},R_p} - \Gamma_{{\rm D},R_p}}
		   {\Gamma_{{\rm A},R_p}} \right| &=
\left| 1 - \frac{\Gamma_{{\rm D},R_p}}{\Gamma_{{\rm A},R_p}}\right| \\
&=
\left| 1 - \left(\frac{\Gamma_{t,s}N_s + 2\Gamma_{t,d}N_d}{N_s + 2N_d}  \cdot 
\frac{N_s + N_d}{\Gamma_{t,s} N_s }\right)\right| \\
&= 
\left|
1 - \frac{(1 + 2 \beta \Gamma_{t,d}/\Gamma_{t,s})(1 + \beta)}{(1+2\beta)}
\right|,
\end{align}
for
\begin{equation}
\beta \equiv N_d/N_s = {\rm BF} \times (1+\gamma_R)^{3/2}.
\end{equation}

For the nominal G2V dwarf case of ${\rm BF}=0.45$ with twin binaries ($\gamma_R 
= 1$) and $\Gamma_{t,d} = \Gamma_{t,s}$ this gives a relative error of $127\%$.
For instance, in the numerical simulation corresponding to 
Fig.~\ref{fig:snr_dist_analytic_v_numeric}, Astronomer A finds $\Gamma_{\rm 
A,R_p} = 0.22$, while Astronomer 
D derives the true (input) occurrence rate of $\Gamma_{\rm D,R_p} = 0.5$.

@@

Write $\Gamma_{{\rm A,\ planets\ of\ }R_p} = \Gamma_{{\rm A,}R_p}$, 
and similarly for ${\rm D}$.
Just as in Model \#1, the answer to ``what is the relative difference between 
the occurrence rates derived by Astronomers D and A for planets of radius 
$R_p$?''
is simpler when we assume that Astronomer A has also derived a 
completeness, which we assume is the same as for Astronomer D in the single 
star case.
So Astronomer A now misclassifies planetary radii, and miscounts the total 
number of stars, but knows his completeness for single stars.
Astronomer D corrects all these errors.
Then
\begin{equation}
	\Gamma_{{\rm A}, R_p} = \frac{N_{\rm det,s}/(f_{s,g} f_{s,c})}{N_s + N_d}.
\end{equation}

The relative difference between the two occurrence rates, $\Delta \equiv 
\left|(\Gamma_{{\rm A},R_p} - \Gamma_{{\rm D},R_p})/\Gamma_{{\rm A},R_p} 
\right|$, is
\begin{align}
	\Delta &=
	\left| 1 - \frac{\Gamma_{{\rm D},R_p}}{\Gamma_{{\rm A},R_p}}\right| \\
	&=
	\left| 1 - \left(\frac{\Gamma_{t,s}N_s + (\Gamma_{t,d1} + 
	\Gamma_{t,d2})N_d}{N_s + 2N_d}  
	\cdot 
	\frac{N_s + N_d}{\Gamma_{t,s} N_s }\right)\right| \\
	&= 
	\left|
	1 - \frac{(1 + \beta (\Gamma_{t,d1}+\Gamma_{t,d2})/\Gamma_{t,s})(1 + 
	\beta)}{(1+2\beta)}
	\right| \\
	&=
	\left| 1 - \left(
	\left[ 1 + \chi \beta \right]
	\cdot 
	\frac{1 + \beta}{1 + 2\beta }
	\right) \right|
\end{align}
for
\begin{equation}
\beta \equiv N_d/N_s, \quad \chi \equiv 
(\Gamma_{t,d1}+\Gamma_{t,d2})/\Gamma_{t,s}.
\end{equation}

In the limit where the $R_p$ planet has the same occurrence rate about every 
host type regardless of its mass, $\chi=2$, and this simplifies to $\Delta = 
\beta$, for a relative error of $59\%$.
This is less than for Model \#1 because of the smaller volume of searchable 
binary stars in Model \#2.

>>>>

Finally, binarity may affect transit surveys through actual astrophysics.
For the relatively rare case [Duquennoy \& Mayor 1991, Raghavan et al 2010] of 
close binaries, dynamical stability imposes a limited range of periods in which 
planets may orbit [e.g., Holman \& Weigert 
1999, Li et al 2017], the effects of which have been observed [Kraus et al 
2016].
Also, the formation rates of planets, and thus their intrinsic occurrence 
rates, may be influenced by properties of circumprimary and/or circumbinary 
disks.
Although the observations are still ongoing, early results seem to indicate 
that once binary companions are $\gtrsim 100\,{\rm AU}$ apart, binarity does 
not decrease planet occurrence rates [CITE], and may even increase them for 
giant planets [Ngo et al 2017].

@@

This is almost simply asking ``what is the relative difference between the 
occurrence rates derived by Astronomers D and A for planets of radius $R_p$?''
However, in the more realistic case, Astronomer A also has derived a 
completeness, which we assume is the same as for Astronomer D in the single 
star case.
So Astronomer A now misclassifies planetary radii, and miscounts the total 
number of stars, but knows his completeness for single stars.
Astronomer D corrects all these errors.
Then
\begin{equation}
\Gamma_{{\rm A}, R_p} = \frac{N_{\rm det,s}/(f_{s,g} f_{s,c})}{N_s + N_d}.
\end{equation}

To express ``how wrong'' our occurrence rate is, we define a correction factor 
$X_\Gamma$ such that the true occurrence rate is the product of the correction 
factor with the measured occurrence rate: $\Gamma_D \equiv \Gamma_A X_\Gamma$.

The correction factor is then
\begin{align}
X_\Gamma &=
\frac{\Gamma_{{\rm D},R_p}}{\Gamma_{{\rm A},R_p}} \\
&=
\frac{\Gamma_{t,s}N_s + 2\Gamma_{t,d}N_d}{N_s + 2N_d}  \cdot 
\frac{N_s + N_d}{\Gamma_{t,s} N_s } \\
&= 
\frac{(1 + \beta)(1 + 2 \beta \Gamma_{t,d}/\Gamma_{t,s})}{(1+2\beta)},
\end{align}
for
\begin{equation}
\beta \equiv N_d/N_s = {\rm BF} \times (1+\gamma_R)^{3/2}.
\end{equation}

For the nominal G2V dwarf case of ${\rm BF}=0.45$ with twin binaries ($\gamma_R 
= 1$) and $\Gamma_{t,d} = \Gamma_{t,s}$ this gives $X_\Gamma = 2.27$.
For instance, in the numerical simulation corresponding to 
Fig.~\ref{fig:snr_dist_analytic_v_numeric}, Astronomer A finds 
$\Gamma_{{\rm A},R_p} = 0.22$, while Astronomer 
D derives the true (input) occurrence rate of $\Gamma_{{\rm D},R_p} = 0.5$.
This gives a relative percentage error of $\delta\Gamma_{{\rm D},R_p} \equiv 
(\Gamma_{\rm A,R_p} - \Gamma_{\rm D,R_p}) / \Gamma_{\rm D,R_p} = -56\%$.

@@

Write $\Gamma_{{\rm A,\ planets\ of\ }R_p} = \Gamma_{{\rm A,}R_p}$, 
and similarly for ${\rm D}$.
Just as in Model \#1, the answer to ``what is the relative difference between 
the occurrence rates derived by Astronomers D and A for planets of radius 
$R_p$?''
is simpler when we assume that Astronomer A has also derived a 
completeness, which we assume is the same as for Astronomer D in the single 
star case.
So Astronomer A now misclassifies planetary radii, and miscounts the total 
number of stars, but somehow knows his completeness for single 
stars. This is possible, because Astronomer A can correctly derive a 
completeness estimate and geometric transit probability for single stars only. 
For Astronomer $A'$ below, including the planets in binary star systems in the 
numerator actually introduces confusion that precludes nice math.
Then
\begin{equation}
	\Gamma_{{\rm A}, R_p} = \frac{N_{\rm det,s}/(f_{s,g} f_{s,c})}{N_s + N_d}.
\end{equation}

The correction factor $X_\Gamma$ is
\begin{align}
X_\Gamma &=
\frac{\Gamma_{{\rm D},R_p}}{\Gamma_{{\rm A},R_p}} \\
&=
\frac{\Gamma_{t,s}N_s + (\Gamma_{t,d1} + 
	\Gamma_{t,d2})N_d}{N_s + 2N_d}  
\cdot 
\frac{N_s + N_d}{\Gamma_{t,s} N_s } \\
&= 
\frac{(1 + \beta (\Gamma_{t,d1}+\Gamma_{t,d2})/\Gamma_{t,s})(1 + 
	\beta)}{(1+2\beta)} 	\\
&=
(1 + \chi \beta )
\cdot 
\frac{1 + \beta}{1 + 2\beta }
\end{align} 
for
\begin{equation}
\beta \equiv N_d/N_s, \quad \chi \equiv 
(\Gamma_{t,d1}+\Gamma_{t,d2})/\Gamma_{t,s}.
\end{equation}

In the limit where the $R_p$ planet has the same occurrence rate about every 
host type regardless of its mass, $\chi=2$, and this simplifies to $X_\Gamma = 
1+\beta$.
Although this is same expression as what we found for Model \#1, the value of 
$\beta$ will be smaller in this model because of the 
smaller volume of searchable binary stars.
